---
title: "FPP Ch 8"
output: html_notebook
---

# Qn 5

```{r, results='hide', message=FALSE}

install.packages("data.table")
install.packages("readr")
install.packages("USgas")
install.packages("readxl")
install.packages("ggdist")

```

```{r}
library(data.table) 
library(fpp3)
library(readr)
library(USgas)
library(scales)
library(readxl)
library(ggdist)

```

```{r}
ec_can <- global_economy %>% 
  filter(Country == "Canada") 

ec_can %>% autoplot(Exports)

can_fit <- ec_can %>%
  model( ETS(Exports ~ error("A") + trend("N") + season("N")))
```

```{r}
can_fc <- can_fit %>%
  forecast(h=10)

can_fc %>% autoplot() +
  autolayer(ec_can, Exports)

# environmentName(environment(accuracy))
accuracy(can_fit)
```

5d) The model with a trend is not significantly better than the model without. likely because there was a significant spike in exports in 2000 and it goes back down to the long term average after.

```{r}
can_fit <- ec_can %>%
  model( 
    ANN = ETS(Exports ~ error("A") + trend("N") + season("N")),
    AAN = ETS(Exports ~ error("A") + trend("A") + season("N"))
  )

accuracy(can_fit)
```

5e) the forecast for AAN with the negative trend seems incorrect. it puts too much weight on the drop since the 2000 spike, instead of the recent points which show a clear increasing trend.

```{r}
can_fc <- can_fit %>% 
  forecast(h=10)

head(can_fc)

can_fc %>% autoplot()

can_fc %>% ggplot(aes(x = Year)) +
  geom_line(aes(y = .mean, color = .model)) +
  geom_line(aes(x = Year, y = Exports), data = ec_can, )
```

5f) Forecast variance of (A,N,N) is $\sigma_h^2 = \sigma^2 [1 + \alpha^2(h-1)]$

```{r}

can_fit %>% components() # gets you the level slope components of ETS
can_fit %>% tidy() # get model coefficients
can_fit %>% fitted() # get only fitted values, this is less useful
can_fit %>% augment() # get exports and fitted values and residuals in the same dataset

ec_can

accuracy(can_fit)

accuracy(can_fit)$RMSE[1]

```

# Qn 6

Forecast the Chinese GDP from the `global_economy` data set using an ETS model. Experiment with the various options in the `ETS()` function to see how much the forecasts change with damped trend, or with a Box-Cox transformation. Try to develop an intuition of what each is doing to the forecasts.

```{r}
china <- global_economy %>% 
  filter(Country == "China")

china %>% autoplot(GDP)

china_fit <- china %>% 
  model(
    "holt" = ETS(GDP ~ error("A") + trend("A") + season("N")),
    "damped" = ETS(GDP ~ error("A") + trend("Ad") + season("N"))
    # "mult_damped" = ETS(GDP ~ error("A") + trend("M") + season("N")) #multiplicative trend does not exist
  )

china_fc <- china_fit %>% 
  forecast(h=20)

china_fc

china_fc %>%
  autoplot() +
  autolayer(china, GDP)

```

Apply box-cox transformation

```{r}
#I don't know what this "features" funciton do but it seems useful

lambda <- china %>%
  features(GDP, features = guerrero) %>%
  pull(lambda_guerrero)

china %>% autoplot(box_cox(GDP, lambda))

china_fit

# supposedly, applying a transofmration within a model will automatically back transform it. How do we confirm?

china_fit2 <- china %>% 
  model(
    "boxcox" = ETS(box_cox(GDP, lambda) ~ error("A") + trend("Ad") + season("N"))
  )

china_fit2 # I don't know how to combine this into the original mable china_fit

china_fit2 %>% 
  forecast(h=10) %>% 
  autoplot() + 
  autolayer(china_fc) +
  autolayer(china, GDP)

#unclear what's going on with this transformation if not dampened. The forecast just shoots off into infinity.
```

# Qn 7

```{r}
aus_production %>% autoplot(Gas)

aus_fit <- aus_production %>% 
  model(
    "AAdM" = ETS(Gas ~ error("A") + trend("Ad") + season("M"))
  )

aus_fit

#why does this take so long to run?
aus_fc <- aus_fit %>% 
  forecast(h = "3 years")

augment(aus_fit)
aus_fc_hilo <- aus_fc %>% 
  hilo() %>% 
  unpack_hilo(`95%`)
aus_fc_hilo # unpack_hilo() will unpack this object

# forecast plot
aus_fc %>% 
  autoplot(aus_production)
gi
# historical actual vs expected
augment(aus_fit) %>% 
  ggplot(aes(x = Quarter))+
  geom_line(aes(y = .fitted), color = "blue") +
  geom_point(data = aus_production, aes(y = Gas))

# add damped trend and compare RMSE
# let the program decide on the best model specification
```

# Qn 8

```{r}
aus_retail
?aus_retail

set.seed(123456789)
myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1))

myseries
```
