---
title: "fpp-ch-10"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

# Load Packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table) 
library(fpp3)
library(readr)
library(USgas)
library(scales)
library(readxl)
library(ggdist)
library(urca)
library(purrr)

```

# Qn 1

```{r}
typeof(LakeHuron)

lake <- LakeHuron |>  
  as_tsibble()

lake |> index()

lake |> autoplot(value)

lake_mod <- lake |> 
  model(
    tslm_piecewise = TSLM(value ~ trend(knots = c(1920))),
    arima = ARIMA(value ~ trend(knots = c(1920)))
  )

lake_mod

lake_mod |> glance()

lake_mod |> 
  select(arima) |> 
  report()
  

lake_mod |> 
  select(arima) |> 
  fitted() |> 
  autoplot(.fitted, color = "blue") +
  autolayer(lake, value)

fc_lake <- lake_mod |> 
  forecast(h = "30 years")

fc_lake |> 
  filter(.model == "arima") |> 
  autoplot(lake)

```

Does the extrapolated linear trend look realistic? I don't think so, there's no reason to assume that there is an upward trend from 1920 onwards. The year to year variation is much larger

Let's look at the residuals.

```{r}
tslm <- lake_mod |> 
  select(tslm_piecewise) 

tslm |> 
  gg_tsresiduals()

tslm |> 
  augment() |> 
  gg_tsdisplay(.resid, plot_type = "partial")

```


# Qn 2

```{r}
souvenirs |> autoplot(Sales)

souv <- souvenirs |> 
  mutate(surffest = month(Month) == 3 & Month >= yearmonth("1988 Mar"))

# https://stackoverflow.com/questions/6558921/boolean-operators-and
# & is vectorized, meaning it will return a vector of trues and falses
# && is not vectorized and will go left to right and examine one by one and short circuit
# & should be used when creating vectors
# && should be used for programming control-flow and is preferred in if clauses
# You should be absolutely certain your vectors are only length 1, such as in cases where they are functions that return only length 1 booleans. You want to use the short forms if the vectors are length possibly >1. So if you're not absolutely sure, you should either check first, or use the short form and then use all and any to reduce it to length one for use in control flow statements, like if

souv

```

Let's do the model now.

```{r}
souv_mod <- souv |> 
  model(
    tslm = TSLM(log(Sales) ~ surffest + trend() + season()) ,
    arima = ARIMA(log(Sales) ~ surffest + trend() + season())
  )

souv_tslm <- souv_mod |> select(tslm) 
souv_tslm |> report()

souv_arima <- souv_mod |> select(arima) 
souv_arima |> report()

```

There is a bit of a difference to the regression coefficients. The trend is still the same, but the regression coefficients can now more properly pick up the signal, instead of trying to fit the ARIMA error - i.e. if the previous month was high, then the next month will also likely be high.

```{r}

souv_new <- new_data(souv, n = 1*12) |> 
  mutate(surffest = (month(Month)==3) & (Month >= yearmonth("1988 Mar")))

souv_new

fc_souv <- souv_mod |> 
  forecast(new_data = souv_new) # you don't have to specify the forecast length if you give it new data 

fc_souv

fc_souv |> autoplot(souv)

fc_souv |> filter(.model == "tslm") |> autoplot(souv)
fc_souv |> filter(.model == "arima") |> autoplot(souv)

```

There doesn't seem to be a significant difference in the forecasts.

I want to try and plot the lower and upper bound in the same plot:

```{r}
fc_souv <- fc_souv |> 
  hilo() |> 
  mutate(
    lower = `95%`$lower,
    upper = `95%`$upper
  )

fc_souv |> head()

fc_souv |> ggplot(aes(x = Month)) +
  geom_line(aes(y = .mean), color = "blue") +
  facet_grid(.model ~ .) +
  geom_line(aes(y = upper), color = "blue", alpha = .25) +
  geom_line(aes(y = lower), color = "blue", alpha = .25) +
  geom_line(data = souv, aes(x = Month, y = Sales)) +
  scale_y_log10()
 
#unpack_hilo does not work with the recent update (to dplyr I think)
  unpack_hilo(`95%`)

# https://stackoverflow.com/questions/74437685/unnest-dist-column-from-fabletoolsforecast-output
#trying to figure out how to unpack the distribution object
# fc_souv[["Sales"]][[1]]
# 
# typeof(fc_souv$"Sales")
# 
# class(fc_souv$Sales)
# fc_souv |> summarise_all(class)

```


Check Residuals

```{r}
souv_tslm |> gg_tsresiduals()

souv_arima |> gg_tsresiduals()
```


# Qn 3

```{r}
vic_elec_daily <- vic_elec |>
  filter(year(Time) == 2014) |>
  index_by(Date = date(Time)) |>
  summarise(
    Demand = sum(Demand)/1e3,
    Temperature = max(Temperature),
    Holiday = any(Holiday)) |>
  mutate(
    Temp2 = I(pmax(Temperature-25,0)),
    Day_Type = case_when(
      Holiday ~ "Holiday",
      wday(Date) %in% 2:6 ~ "Weekday",
      TRUE ~ "Weekend"))

head(vic_elec_daily)

```

```{r}
vic_elec_daily |> autoplot(Demand)
```

```{r}
vic_elec_daily |> 
  ggplot(aes(x = Temperature, y = Demand, color = Day_Type)) +
  geom_point()
```

```{r}
elec_model <- vic_elec_daily |> 
  model(ARIMA(Demand ~ Temperature + Temp2 + (Day_Type == "Weekday")))

elec_model |> gg_tsresiduals()

```
We can optimize the choice of by first doing a hand selection based on the Temp*Demand graph. Then, we can try a few different knots and see which one performs the best on the testing.

#Qn 4

```{r}
head(aus_accommodation)

aus_accommodation |> 
  distinct(State)

aus_accommodation |> autoplot(CPI)

CPI_1 <- aus_accommodation |> 
  pull(CPI) |> 
  min()

aus <- aus_accommodation |> 
  mutate(
    CPI_index = CPI / CPI_1,
    Takings_CPI = Takings / CPI_index
  )

aus |> 
  filter(State == "New South Wales") |> 
  ggplot(aes(x = Date)) +
  geom_line(aes(y = Takings)) + 
  geom_line(aes(y = Takings_CPI), color = "blue")

index(aus)
key(aus)

```


```{r}
mod <- aus |> model(
  arima_model = ARIMA(Takings_CPI ~ trend(knots = yearquarter(c("2008 Q1"))) + season()) 
)

mod
mod |> glance() # why is no information showing up for the glance?
mod |> tidy()
mod |> filter(State == "New South Wales") |> report()

mod |> forecast(h=5) |> 
  autoplot()

```

```{r}

for(i in 1:nrow(mod)) {
  print(i)
  mod[i,] |>  
    gg_tsresiduals() |> 
    print()
}

```

I am trying to functionally loop over and print the graphs. Can't seem to figure it out. The closes I've gotten is the code in the bottom.

Maybe also read through ch 11 to see if there are any thoughts there


```{r}

# NOT RUN {
require(stats)
warpbreaks

by(warpbreaks[, 1:2], warpbreaks[,"tension"], summary)
by(warpbreaks[, 1],   warpbreaks[, -1],       summary)
by(warpbreaks, warpbreaks[,"tension"],
   function(x) lm(breaks ~ wool, data = x))

## now suppose we want to extract the coefficients by group
tmp <- with(warpbreaks,
            by(warpbreaks, tension,
               function(x) lm(breaks ~ wool, data = x)))
sapply(tmp, coef)
# }

test <- mod |> split( mod$State) 
test |> class()
test[1] |> as_mable() |> class()

mod

key(mod)

###############################

print_info  <- function(obj, in_key){
  obj |> class() |> print()
  obj |> head() |> print()
  obj |> as_mable() |> class()
  # obj |> 
  #   mutate(key = in_key) |> 
  #   setkey(key) |> 
  #   as_mable() |> 
  #   head() |> 
  #   print()
 
}

mod |> group_by(row_number()) |>  group_walk(print_info)
#check whether you have a ls_modl as your variables - so print your var names and var types and see what happens

mtcars

mtcars %>%
  group_by(cyl) %>%
  group_map(~ head(.x, 2))

mtcars %>%
  group_map(~ head(.x, 5))

```

```{r}

mod |> class()

test <- mod |> 
  t() |> 
  as.data.frame(col.names = mod[[State]]) 

test

test |> length()

test[1] |> class()

mod
map(mod, gg_tsresiduals())

print_residuals <- function(in_mable) {
  #print(class(in_mable))
  in_mable |> gg_tsresiduals()
}

mod |> 
  mutate()

walk(mod$arima_model, print_residuals)


mod |> class()
mod |> pull(arima_model) |> class()
mod$arima_model |> class()

mod

mod
mod[1,] |> gg_tsresiduals()
mod[1,2]

mod |> t() |> t() |> class()

test <- mod |> group_by(State) |> group_split()

test |> class()

test[1] |> as_mable() |> class()

print_class <- function(obj){
  obj |> 
    class() |> 
    print()
}

walk(mod, class)

```






