---
title: "fpp-cheatsheet"
output: html_document
date: "2023-03-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Fabletools to apply to models 

You can look at the fable, fabletools, tsibble pages for more information.

```{r}

can_fit %>% components() # gets you the level slope components of ETS
can_fit |> report() #gives you the model printout
can_fit %>% tidy() # get model coefficients
can_fit %>% fitted() # get only fitted values, this is less useful
can_fit %>% augment() # get exports and fitted values and residuals in the same dataset
can_fit |> accuracy() # gives you RMSE, etc.
ec_can

```

# Typeof vs Class


votes

The class() is used to define/identify what "type" an object is from the point of view of object-oriented programming in R. So for

> x <- 1:3
> class(x)
[1] "integer"
any generic function that has an "integer" method will be used.

typeof() gives the "type" of object from R's point of view, whilst mode() gives the "type" of object from the point of view of Becker, Chambers & Wilks (1988). The latter may be more compatible with other S implementations according to the R Language Definition manual.










